\documentstyle[11pt]{article}
\setlength{\textheight}{9in}
\setlength{\topmargin}{0in}
\setlength{\baselineskip}{0.1in}
\setlength{\textwidth}{6.3in} 
\setlength{\oddsidemargin}{0in}  

\def\cf{{\it cf.}}
\def\eg{{\it e.g.}}
\def\ie{{\it i.e.}}
\def\etal{{\it et al.}}
\def\etc{{\it etc}}
\def\ni{\noindent}
\def\go{\mathrel{\raise.3ex\hbox{$>$}\mkern-14mu\lower0.6ex\hbox{$\sim$}}}
\def\lo{\mathrel{\raise.3ex\hbox{$<$}\mkern-14mu\lower0.6ex\hbox{$\sim$}}}
\begin{document}
\parindent=0pt
\section*{Biographical Sketch}
\subsection*{Roger David Blandford}
Luke Blossom Professor in the School of Humanities and Sciences

Professor at the Stanford Linear Accelerator Center.
\subsection*{Professional Preparation}
B.A. (Theoretical Physics), Cambridge, 1970

M.A. (Astronomy), Cambridge, 1973

Ph.D. (Applied Mathematics), Cambridge, 1973

\subsection*{Previous Appointments}
2003 - 2013 Pehong and Adele Chen Director Kavli Institute for Particle Astrophysics and Cosmology

1992 - 1995 Executive Officer for Astronomy, Caltech

1989 - 2004 Richard Chace Tolman Professor Theoretical  Astrophysics, Caltech 

1979 - 1989 Professor, Caltech

1976 - 1979 Assistant Professor, Caltech

1973 - 1976 Research Fellow, St. John's College, Cambridge

1972 - 1973 Charles Kingsley Bye Fellow, Magdalene College, Cambridge 

\subsection*{Societies}
Fellow Royal Astronomical Society (1974)

Fellow of Royal Society (1989)Fellow of American Academy of Arts and Sciences (1993)

Member National Academy of Sciences (2005)

Fellow of American Physical Society (2010)

\subsection*{Synergistic Activities}
Chair Astronomy and Astrophysics Decadal Survey (2010)

Chair: NSF Astronomy Senior Review Committee (2007)

Chair: American Astronomical Society, High Energy Astrophysics Division (2004-5)

Annual Reviews Astronomy and Astrophysics Editor (2002-4), Managing Editor (2005-10)

Instigator and co-organiser of First NSF-ITP Workshop for High School Physics Teachers (2001)

Instigator of and original lecturer for Caltech Core Curriculum course for freshman not intending to study astronomy:  ``The Evolving Universe'' (2000)

Co-author (with K. Thorne) of graduate text: ``Modern Classical Physics (2015, in press)

\vfill\eject

\subsection*{Selected Recent Publications}
Collett, T., et al. 2013 MNRAS 418 3183

Greene, Z. et al. 2013 ApJ 768

Hezaveh, Y., Marshall, P. \& Blandford, R. 2014 ApJ submitted.

Morganson, E. et al 2010 MNRAS 406 2452

Oguri, M. et al. 2013 MNRAS 429 482

Ota, N.  et al 2102A\&A 547 A77

Suyu, S. 2010, ApJ 711 201

Suyu, S. et al. 2014 ApJ 788 L35

Thorne, K. \& Blandford, R. 2015 Modern Classical Physics Princeton University Press

von der Linden, A. et al. 2014 MNRAS 439 2

\subsection*{Recent Collaborators}
E. Agol, S. Allen,  A. Broderick, M. Begelman, P. Coppi, C. Fassnacht,  L. Koopmans,  M. Lyutikov, M. Malkan,  D. Rappetti, Y. Rathore, S. Suyu. T. Treu, M. Bietenholz, R. Buehler, P. Eggleton, S. Funk
\subsection*{Recent Graduate Students}
P. Simeon, M. Amin, C. Zheng, E. Morganson, J. Dai, Y. Yuan, R. Anantua.
\subsection*{Recent Postdoctoral Scholars}
P. Marshall, M. Oguri, L. Stawarz, W.
\subsection*{PhD Adviser} 	
M. Rees
\end{document}
